Computational hysteresis:\\ inter-task effects in human computation

### 2014-11-16 docs/publications/2014.09_SCIENCE/new_man.tex
further analysis, plots, manuscript work

People still outperform sophisticated artificial 
intelligince systems in many tasks requiring general knowledge, 
qualitative judgment, and visuo-spatial reasoning. Recently, microtask 
platforms enable the programmer to import human judgment directly into 
algorithms: programs can post tasks, to be completed by users of such 
platforms in near real-time.
Some computer scientists consider microtask platforms to be a fundamentally
new computing architecture, but psychological effects can introduce systematic
bias. 
Here we investigate the effects that earlier tasks have on the results of 
subsequent ones, which we call \textit{inter-task effects}.
In a canonical image labeling task, we found that inter-task effects are
very strong---stronger than overtly framing tasks in terms of 
particular research goals.  However, in some cases, inter-task effects can
be leveraged to elicit more nuanced responses.
To study inter-task effects we introduce a framework that characterizes HPU 
bias in an application-independant way, based on its potential to alter an 
algorithm's path of execution.  Our findings suggest that careful consideration
should be given to inter-task effects when designing crowdsourcing studies.

%microtask worker performs a series of tasks.  We find that early tasks can
%have a strong effect on later tasks, and that these effects can be stronger
%than those brought about by framing.
%
%course of human computation algorithms.
%application-independant quantity that measures the
%severity of priming.  We apply this framework to measure 
%
%in many
%tasks, especially ones requiring general knowledge about the world, qualitative
%judgment, and visuo-spatial reasoning.  Recently, microtask web platforms have 
%emerged. One can upload a batch of tasks to be completed by users who are paid 
%for their efforts.
%Though simple on the surface, these platforms dramatically enlarge the
%programmers toolkit: in effect they allow the programmer to import human 
%judgment into algorithms.  Some computer scientists consider
%such platforms to be a new computing architecture.  But people are very 
%different from electronic processors.  For example, context affects how a 
%person will respond to a prompt, and this complicates the formal treatment of 
%human computation.  To approach this issue, we introduce a framework 
%that quantifies the severity of priming in an application-independant way.
%We apply this technique to reveal an unforseen source of bias: the bias 
%induced by one task on another.  Surprisingly we find that 
%\textit{inter-task effects} introduce more bias than \textit{framing}, which
%has received more attention.
%
%more well-known 
%\textit{framing}, a more well-studied form of priming.
%
%Microtask platforms offer a fast, flexble, and inexpensive source of labour for
%clerical tasks \cite{Finnerty2013, chandler2013breaking, Berinsky2012351} and 
%recently have been adopted by researchers to augment datasets and solicit 
%participants \cite{paolacci2010running, Berinsky2012351, chandler2013breaking}.
%Computer scientists have described such platforms as a new computing 
%architecture, and seek to characterize “human co-processing units” (HPUs) in 
%analogy to CPUs and GPUs \cite{5543192}. However, it is known that people are 
%susceptible to 
%priming effects\cite{No2007,Swaab200299,Gopher2000308,sohn2001task,Ghuman17062008,BJOP1796,BJOP1826,Gass1999549}, which stands to complicate the HPU 
%perspective.  Here, we introduce a framework to recast this psychological 
%phenomenon as a kind of computational hysteresis. 
%We apply this framework to study \textit{inter-task effects} that arise when 
%microtask 
%workers perform a sequence of tasks. Surprisingly, earlier tasks can have a 
%very strong influence on workers during later tasks---stronger than overtly
%framing the tasks as serving different research goals.   
%This result has major implications for the design of crowdsourcing initiatives.
%The framework we introduce provides a general measure of priming severity, 
%and, in combination with machine learning tools, enables measurement without
%knowing ahead of time how priming will affect output.
\end{sciabstract}

@@

These platforms make it possible to seemlessly integrate human and 
machine computation.  Researchers coined the term HPU 
(Human co-Processing Unit) in analogy to CPUs and GPUs \cite{5543192}.  
Attempts have been made to establish an HPU instruction set.  For example,
the X library provides function calls such as 
\texttt{X.vote()}, and \texttt{X.map()}
 \cite{little2010turkit,minder2011crowdlang,minder2012crowdlang,kittur2011crowdforge}.

Of course, there are obvious differences between HPUs and their silicone
counterparts.  When facing the same task, people do not generally
respond in the same way.  Common sense would dictate that a certain amount
of variability in responses made by people is inescapable.  But an important 
part of the variation is bound to be systematic, due to psychological or 
demographic factors.

Here, we highlight an important source of bias with
serious implications for any crowdsourcing initiative. 
It is known that people are susceptible to priming effects 
\cite{BJOP1796,No2007,beller1971priming}, and, in particular, task-repetition 
effects \cite{Gass1999549,sohn2001task}.  Thus, a worker's response during
one task may depend in part on previous ones.  Such \textit{inter-task} effects
would amount to a kind of \textit{hysteresis}, meaning that HPU outputs are not
only a function of the current input, but also of the history of inputs.

@@

We make two contributions towards understanding inter-task effects,
and systematic bias in HPUs in general.  First, we present a framework for 
quantifying bias in HPU outputs in terms of their potential impact 
algorithms.  This yields an approach for measuring the severity of bias 
that removes the need to understand the underlying psychological or 
demographic factors at play.  Second, we apply our framework to measure the
severity of inter-task effects on a canonical microtask: image labeling.

Using the  Amazon Mechanical Turk (Mturk) platform, we solicited workers to 
perform a series of image-labeling tasks.  We chose image labeling because it 
is one of the most common kinds of microtasks
\cite{chandler2013breaking,Berinsky2012351,Finnerty2013,paolacci2010running}, 
and it seems likely that this will remain the case in the long term because
of applications in computer vision research \cite{5543192}.
%In these tasks, workers labeled images featuring iconic cultural scenes,
%food, table-settings, and various utensils and cookware. 
We divided the tasks into an \textit{initial} and \textit{test} set, 
but, crucially, there was no distinction between these sets from the
perspective of the worker.  We varied the initial tasks, while 
keeping the test set the same, to analyze the effect of initial tasks on 
the labels provided during the test tasks.

As a point of comparison, we included treatments in which we induce 
\textit{framing}.
Before the workers labeled images, we indicated that the work 
was funded by a research lab with a particular area of interest, to see the 
influence this has on the workers’ subsequent labels. 
Surprisingly, inter-task effects were stronger than framing, in some cases 
about twice as strong. 
Our results demonstrate that initial tasks can change workers' focus in later 
tasks.  We also found that inter-task effects altered the level of specificity 
of worker's labels.  Our findings suggest that careful consideration should be 
given to the bundling of tasks when designing a study using a microtask 
platform.

\paragraph*{Computational hysteresis.}
Before describing our experiments, we dedicate some space to discussing
how we characterize and measure priming in HPU outputs.  Since HPU outputs are
stochastic, determining the effect of priming on an HPU population
amounts to determining the resulting change to the HPU population's 
distribution of outputs.

To refine this, we distinguish between methods that determine \textit{whether}
two random sources have the same distribution 
(e.g. Pearson's chi squared test), from those that determine
\textit{the extent} of that difference (e.g. one of the many measures of 
statistical divergence).  We are interested in measuring the extent of
difference because our practical concern is the potential for such
differences to alter the course of an algorithm.

To measure the effect of some priming treatment, we must of course draw a 
comparison to some standard population.  In many cases, there may be an 
obvious comparison between a primed and unprimed population.  However, we 
refrain from proposing a canonical neutrally-primed HPU population; what is 
considered ``unprimed'' in one scenario may be ``primed'' in another. 
Instead, we speak of two HPU populations being \textit{differently primed}.

We present three alternative, equivalent definitions for the 
\textit{difference in priming} of two HPU populations, $\mathcal{P}$, and 
$\mathcal{Q}$, which we denote $\theta(P,Q)$.  To set up the definitions,
we will assume that $\mathcal{P}$ and $\mathcal{Q}$ 
were sampled randomly from the same initial population $\mathcal{R}$, and 
then were exposed to (possibly) different priming treatments. Then, the HPUs 
in $\mathcal{P}$ and $\mathcal{Q}$ perform some task, leading to popualtions 
of work-products, $P$ and $Q$.

The first definition provides the motivation for our choice of divergence
measure, because it directly measures the proportion of the time that a
binary decision will be changed due to priming effects:

\begin{mydef}
	\label{def:bias}
	{\upshape Maximum Bias:}
	Let $\mathcal{D}$ be a binary decision function, which takes a 
	work-product
	$r \in P \cup Q$ as input, and outputs a bit $\mathcal{D}(r)=b$, where 
	$b \in \{0,1\}$. Denote the \emph{bias} of $\mathcal{D}$, due to the
	different priming of $\mathcal{P}$ and $\mathcal{Q}$ by
	$\theta_\mathcal{D}(P,Q)$ and define it to be the difference 
	in the expected output of $\mathcal{D}$ when acting on work-products from 
	$P$ and $Q$:
	$$
	\theta_\mathcal{D}(\mathcal{P},\mathcal{Q}) = 
		\left| 
			\mathrm{E}\left\{ \mathcal{D}(P) \right\} 
			- \mathrm{E}\left\{ \mathcal{D}(Q) \right\} 
		\right|,
	$$
	where $p \in P$, $q \in Q$.  Then, define the difference in priming 
	between $\mathcal{P}$ and $\mathcal{Q}$ to be the worst-case bias, i.e. 
	the bias of the most susceptible decision:
	$$
	\theta(\mathcal{P},\mathcal{Q}) = 
		\sup_\mathcal{D} \theta_\mathcal{D}(P,Q)
	$$
\end{mydef}
If $\mathcal{D}$ is the condition of an \texttt{if}-statement, then the 
definition above is the change in the fraction of the time that a given 
path of execution is followed due to HPU priming.

While this is precisely what we want to measure, it is not at all clear how
to go about doing it.  The second formulation provides a practical connection
by rephrasing priming in terms of its distinguishability:

\begin{mydef}
	{\upshape Distinguishability:}
	Let $\mathcal{D}$ be a distinguisher algorithm, which takes work-product 
	$r \in P \cup Q$ as input, and outputs a \emph{guess} 
	$\mathcal{D}(r) \in \{0, 1\}$. Now define a validation test 
	$V(\mathcal{D}, R_0, R_1)$ as follows. A bit is drawn uniformly at random, 
	$b \in_R \{0, 1\}$, and then a work-product $r$ is chosen uniformly at 
	random from one of the populations, decided by the bit: $r \in_R R_b$. 
	Input $r$ into $\mathcal{D}$, and if $\mathcal{D}(r) = b$, we say 
	$\mathcal{D}$ \emph{guessed correctly}, and the value of $V$ is 1, 
	otherwise $V = 0$.

	A distinguisher that guesses randomly will still be correct half of the 
	time. Therefore, denote the performance of $\mathcal{D}$ by 
	$\eta_\mathcal{D}(P, Q)$, and let
	$$
		\eta_\mathcal{D}(P,Q) = 2\cdot \mathrm{E}\{V(\mathcal{D},P,Q)\} - 1.
	$$
	Then define the difference in priming between $\mathcal{P}$ and 
	$\mathcal{Q}$ to be the performance of the best possible distinguisher:
	$$ 
		\theta(P,Q) = \sup_\mathcal{D} \eta_\mathcal{D}(P,Q)
	$$
	\label{def:dist}
\end{mydef}

\textbf{Definition 2} relates the difference in priming to the 
\textit{optimal Bayes classifier} for determining whether a given HPU output
came from $P$ or $Q$.  This means that we can obtain a lower bound on
$\theta(P,Q)$ based on the performance of an actual classifier, which provides
recourse when purely statistical methods fail.

The statistical divergence measure underlying the previous definitions is the
divergence measure known as the ``total variational distance'' or 
$l_1$-distance:
\begin{mydef}
	\label{def:l1}
	{\upshape Total variational distance $\theta_{l_1}$:}
	Define the difference in priming between $\mathcal{P}$ and $\mathcal{Q}$ 
	to be equal to the total variational distance between the work-products 
	$P$ and $Q$:
	$$
	\theta_{l_1}
	= \sup_{A \subset \mathcal{X}} \left| \sum_{x \in A} f_P(x) - f_Q(x)\right|
	= \frac{1}{2} \sum_{x \in \mathcal{X}} \left| f_P(x) - f_Q(x) \right|,
	$$
	where $f_P(x)$ denotes the probability of observing x when sampling from P.
\end{mydef}
The equivalent formulation for $\theta(P,Q)$ in 
\textbf{definition \ref{def:l1}} provides a grounding in a fundamental
measure of statistical divergence, and a connection to information theory.  
The equivalence of the three definitions
follows from the known fact that no algorithm can have output that differs
on inputs $P$ and $Q$ more than $l_1 \times 100\%$ of the time.  We provide
proofs in the supplementary material.

The difficulty of measuring $\theta(P,Q)$ depends on what the universe of
possible outputs are, as well as how $P$ and $Q$ are actually distributed,
both of which are mainly determined by the task that the HPUs are asked to 
perform.  In the case where workers label an image with a free-text input, 
the number of different possible labels can be in the thousands or more.  If
the support, $\mathcal{X} = P \cup Q$, is large compared to the number of 
HPU outputs sampled, so that many labels appear only once (which in any case
is expected from Zipv's law), then estimating $\theta$ directly from the
observed frequencies of labels is difficult.

One approach to estimating the $l_1$-distance from samples, which appears
reasonable on the surface, is to use the observed frequencies of each output
$x$ from $P$ and $Q$ to estimate $f_P(x)$ and $f_Q(x)$, and calculate 
$\theta(P,Q)$ directly from \textbf{definition 3}.  This is the so-called naive
approach to estimating $l_1$-distances, and is aptly named because it tends
to produce very inflated values.  More recent techniques have for estimating
$l_1$ have been suggested by Valiant \cite{val-thesis}, and Batu 
\cite{batu2000testing,batu2013testing}.  We reserve the discission of these
techniques to the supplementary material; they do not turn out to be useful
because they do not provide confidence intervals.  Thus our only recourse is
to test the performance of classifiers at distinquishing $p \in P$ from 
$q \in Q$, which allows us to estimate a lower bound on $\theta(P,Q)$.  This 
turns out to be sufficient, and has the advantage that, so long as they are
learnable, it is not necessary to understand exactly how $P$ and $Q$ differ.

\paragraph{Experimental setup.}
We performed two experiments, soliciting respectively 900 and 2300 MTurk 
workers to label images depicting iconic cultural scenes, meals, ingredients, 
table-settings, and various utensils and cookware. In these experiments,
workers were randomly assigned to treatments, and depending on this 
assignment, were exposed to different framing primes, given a different 
set of initial images, and shown a different permutation of the test images. 
For both experiments, workers were first shown brief instructions.  
The framing prime was then shown, unless the given 
treatment involved no framing.  Workers then performed a series of 
image-labeling tasks and had to provide five labels per image. 
For the purpose of analysis, we denote the first five images as the 
\textit{initial set}, and the last five as the \textit{test set}, but no
distingction was made from
the perspective of the worker.
Table 1 sumarizes the various experimental treatments for both 
experiments.

\begin{table}[t]
\centering
	\begin{tabular}{ l  l  l }
		\hline                       
		Treatment & Framing prime & Initial image set	\\ 
		\hline                       
		\textit{Experiment 1}\\
		$\textsc{ambg}$ & None & Ambiguous\\
		$\textsc{img:cult}$ & None & Cultural\\
		$\textsc{img:ingr}$ & None & Ingredients\\
		$\textsc{frm:cult}$ & Cultural & Ambiguous\\
		$\textsc{frm:ingr}$ & Ingredients & Ambiguous\\
		\hline  
		\textit{Experiment 2}\\
		$\textsc{img:food}$ & None & Food\\
		$\textsc{img:obj}$ & None & Objects\\
		$\textsc{frm:food}$ & Food & None\\
		$\textsc{frm:obj}$ & Objects & None\\
		\hline  
	\end{tabular}

	\caption{ \footnotesize{ 
		Workers were uniformly randomly assigned to one of the 
		treatments listed above. 
		The full funder names used were 
		Cultural: ``The Global Foundation for Cultural Recognition'' and 
		Ingredients: ``The National Foundation for Nutritional Awareness''.  
		The ambiguous, cultural, and ingredients initial image sets are shown 
		in \textbf{Figs. 2}, \textbf{3}, and \textbf{4}.
	}}
	\label{table:1}
\end{table}

@@

\paragraph{The strength of inter-task effects.}
\begin{figure}
	\centering
	\includegraphics[scale=1]{figs/theta.pdf}
	\caption{
		Difference in priming observed for various pairs of worker populations.
		The pairs of populations differed in terms of the initial tasks they
		performed (bars labeled by \textit{inter-t.}) or in terms of the framing
		to which they were exposed (bars labeled by \textit{frame}; see table 1 for
		details about the treatments).  The difference in priming was 
		measured based on the accuracy of a naive Bayes classifier in 
		resolving which treatment workers belonged to, as described in the 
		text.  
	}
	\label{fig:theta}
\end{figure}

%\paragraph{Dynamics of inter-task effects}
%\begin{figure}
%	\centering
%	\includegraphics[scale=0.7]{figs/theta_by_pos_img.pdf}
%	\caption{
%		Difference in priming observed for various pairs of worker populations.
%		The pairs of populations differed in terms of the initial tasks they
%		performed (bars labeled by ``inter-t.'') or in terms of the framing
%		to which they were exposed (bars labeled by ``frame''; see table 1 for
%		details about the treatments).  The difference in priming was 
%		measured based on the accuracy of a naive Bayes classifier in 
%		resolving which treatment workers belonged to, as described in the 
%		text.  
%	}
%	\label{fig:theta_pos_img}
%\end{figure}

We analyzed the overall strength of inter-task effects and framing, by looking
at pairs of treatments that performed the same test tasks, but had done
different initial tasks or had been exposed to different framing.
For each such pair, we assessed the performance of classifier
in predicting which treatment workers belonged to, based on the labels they
provided during the test tasks (see fig 2).  
Based on \textbf{Definition 2}, this is a lower bound on the amount 
bias introduced by the initial tasks and framing.

In experiment \textit{inter-t.1}, the test tasks required workers to
label relatively cluttered images of table settings, which had various items
of food, utensils, dishware, and other objects natural to such settings.
For one treatment, the images in the initial tasks consisted exclusively of 
food, and for the other, they consisted exclusively of non-food objects, 
particularly cookware and utensils.  We saw a strong effect of priming, 
based on the classifier performance (see fig. 2).

In a related experiment, \textit{frame1}, workers were given the same test tasks
as in \textit{inter-t.1}, but were either told 
that ``The purpose of this study is to understand the viusual perception of
Food and Ingredients'' or ``\ldots of Objects and Tools''.  This overt framing
of the task in terms of a specific research objective had remarkably milder
effects on workers than the initial tasks (see fig. 2).

We performed another set of experiments based on a different set of test 
tasks, and where the initial tasks and framing were directed along a different
conceptual axis.  In \textit{inter-t.2} and \textit{frame2}, the images in the test 
tasks displayed prepared meals from a variety of ethnic backgrounds.  We then
designed the framing and initial tasks to focus either on food or cultural 
elements.  In the previous experiments, the primes were always focused on 
concrete ``highly imageable'' concepts (food and objects), these experiments
mixed a concrete concept (food) with an abstract concept (culture).
The initial tasks in \textit{inter-t.2} either depicted prepared meals (but having
ambiguous cultural association), or iconic cultural scenes (including music, 
dance, and sport).  Workers in \textit{frame2} were told that the ``This study
was proudly funded by the National Foundation for Nutritional Awareness'' 
or the ``\ldots Global Foundation for Cultural Recognition''.

The results for \textit{inter-t.2} and \textit{frame2} confirmed that 
inter-task effects are much stronger than framing (see fig. 2).  In fact, 
intertask effects were even stronger in this experiment than the previous.

\paragraph{Sufficiency of framing.} Given just how much stronger
inter-task effects were, the reader may wonder whether the framing was 
in some sense ``sufficient''.  
We designed our framing tasks based on
work that had previously studied framing effects in micro-task settings.  
Prior studies have shown that framing the purpose of a task, or indeed, 
simply changing a single, seemingly non-critical word in a task description, 
has significant effects on worker's resonses \cite{thibodeau2013natural}.  Perhaps more importantly, the 
framing treatments were included to provide a reasonable basis for comparison,
and it is clear from our results that the prior tasks performed by a 
worker are much stranger than the effects induced by descriptions of a task's
purpose, or the requester's mandate.

Nevertheless, we included one more experiment, \textit{frame*3}, to test the 
effect of framing when the worker is required to \textit{echo} the framing 
statement.  \textit{frame*3} was based on \textit{frame2}, but the worker
had to indicate, using a menu option, what the purpose of the study was,
before they could proceed.  In one sense, this seems to go beyond what one
might mean by ``framing''.  By forcing the worker to echo the frame, we provide
a clear indication that the frame is a particularly important detail, 
elevating its importance above even the instructions that had been provided
beforehand.

The framing effect in \textit{frame*3} was quite strong, being on par with the
inter-task effects we had observed (see fig. 2).  We are not sure whether
the \textit{frame*3} should be considered framing,
or whether it should be considered as an instruction, or perhaps as a 
screening-task.  Regardless, our results show that worker's responses are 
profoundly affected by recently completed tasks.

\paragraph{Dynamics of intertask effects.} 
In the results shown above, we considered the aggregate effect that the
first five initial tasks had on the last five test tasks.  But it would stand
to reason that the first test task might be
most strongly influenced, and that, as the worker proceeds, the intervening 
test tasks would tend to wash out the priming induced by the initial tasks.

>>>>

Microtask platforms combine the efforts of large numbers 
of people to perform tasks that are difficult to automate with computers 
alone.  These platforms support near-real-time task completion, mimicking a
compute server, and fundamentally shift the role of the human from being a
consumer to being a producer of computing resources.  To realize its full
potential, we must understand how human computation differs from machine 
computation, and in particular address the biases known to be inherent in 
human cognition.  Here we show that when a worker completes a series of 
microtasks, the responses in later tasks are strongly influenced by the 
content of earlier ones.  Our findings are not solely cautionary. We find
evidence that these \textit{inter-task effects} can be
harnessed to improve the quality of responses, and help to achieve 
expert-level judgments, to increase the scope and reliability of citizen 
science and crowdsourcing initiatives.
\end{sciabstract}

@@

These platforms make it possible to tightly integrate human and 
machine computation.  Researchers coined the term HPU 
(Human co-Processing Unit) in analogy to CPUs and GPUs \cite{5543192}.  
Ongoing research pursues the establishment of an HPU instruction set
including basic operators like \texttt{vote}, \texttt{subdivide}, \texttt{map},
\texttt{reduce}, \texttt{sentiment}, and \texttt{image\_tag}.  As was 
done with GPU programming, libraries are being developped to give the 
programmer more seamless access to HPU resources 
 \cite{little2010turkit,minder2011crowdlang,minder2012crowdlang,kittur2011crowdforge}.  The availability of human computing platforms makes whole new classes
of problems tractible for the programmer.  For example ***.

But the ability to program with high-level human instructions comes with a 
cost: HPU output is inherently noisy.  Part of this noise could be 
explained by the random differences between workers.  Meanwhile, systematic 
cognitive biases also play a role.  
The phenomenon we attend to here represents yet
another dimension of HPU variance, and arises from the fact that a worker's 
performance changes as a function of the content of previous tasks.

@@

It is known that people are susceptible to priming effects 
\cite{BJOP1796,No2007,beller1971priming}, and, in particular, task-repetition 
effects \cite{Gass1999549,sohn2001task}.  Thus, a worker's response during
one task may depend in part on previous ones.  Such \textit{inter-task} effects
would amount to a kind of \textit{hysteresis}, meaning that HPU outputs are not
only a function of the current input, but also of the history of inputs, with
important implications for crowdsourcing initiatives.

Here we describe a series of experiments using the Amazon Mechanical
Turk (MTurk) microtask platform that studied inter-task effects
for image-labeling microtasks.  Image-labeling tasks are among the most 
common kinds of microtasks
\cite{chandler2013breaking,Berinsky2012351,Finnerty2013,paolacci2010running}, 
and it seems likely that this will remain the case in the long term because
they provide a source of ground truth in computer vision research 
\cite{5543192}.  
During the experiments, workers performed a series of image labeling tasks, 
and the first five images
they labeled were varied depending on the experimental group to which they
were assigned.  We studied the effects that changing the initial tasks 
had on the labels provided in the final tasks.

As a point of comparison, we included treatments in which we induced
\textit{framing}.
In the framing treatments, before workers begin the tasks, we told them that 
the study was funded by a lab with a particular area of interest, for example,
in the visual perception of food in one case, or the perception of tools in
another.

Surprisingly, inter-task effects were much stronger than framing.  
Our results demonstrate that initial tasks can change 
workers' focus in later tasks.  We also found that inter-task effects altered 
the level of specificity, and the diversity of vocabulary that workers used,
in ways that were not reproduced using framing.
Contrary to our initial expectations, our findings suggest that by carefully
designing workflows, the HPU programmer can leverage inter-task effects to
direct focus, tune specificity, and activate worker expressivity.  Our 
findings carry two lessons for the task designer: inter-task
effects can severely bias HPU output if HPU hysteresis is not taken into 
account, but at the same time, inter-task effects provide another technique
for improving HPU output and may help elicit expert-level annotation.

\paragraph*{Measuring computational hysteresis.}
In setting out to measure the computational hysteresis induced by priming, 
it is important to remember that we are only interested in the practical 
impact of priming on the execution of HPU algorithms.  
The distinction between measuring priming and measuring computational 
hysteresis is subtle, but it is crucial to ensure that our findings are of 
practical significance to crowdsourcing applications.

This distinction dictates our choice of tools for detecting HPU hysteresis.
A common practice for demonstrating that one population is different from
another (e.g., HPUs exposed to one prime vs another), is to use a 
$\chi^2$ test. This would suffice to demonstrate \textit{whether} priming of 
HPUs had occured, however it says nothing about the \textit{extent} of 
priming, and in particular, does not demonstrate the extent to which primed 
HPUs would execute an algorithm differently.

We opt instead to measure HPU hysteresis using
machine learning algorithms (classifiers), that have been trained to 
distinguish HPUs that were subjected to one prime, from those subjected to 
another.  If a classifier
can distinguish HPUs subjected to different primes, then it serves as 
a kind of certificate, proving 
(up to statistical confidence) that there exists 
an algorithm whose execution is affected by the priming of individual HPUs
(namely, the classifier itself).  This demonstrates that, for some algorithms,
using HPUs primed one way vs another will lead an algorithm to follow a 
different branch of execution with a certain frequency.
Statistically, our approach is grounded in the 
statistical divergence metric called L1-distance 
(alternatively \textit{total variational distance}), via the following 
inequality:
\begin{equation}
	D_\mathrm{L1} \geq 2 \eta - 1, 
	\label{l1}
\end{equation}
where $D_\mathrm{L1}$ is the L1-distance, and $\eta$ is the accuracy of
any binary classifier during validation (provided mild conditions are met). 
Since a classifer's performance estimates
a lower bound on the L1 distance, we shall denote 
$D_\mathrm{L1}^-\equiv 2\eta-1$.
The value of $D_\mathrm{L1}$ ranges from 0 to 1; when $D_\mathrm{L1} = 0$ HPUs 
shown one prime or another are 
indistinguishable, wheras if $D_\mathrm{L1} = 1$ priming is so complete that 
differently primed HPUs can be distinguished perfectly.
The
interested reader will find further discussion of the statistical properties
of our approach and proofs of the above claims in the supplementary material.

\paragraph{Experimental setup.}
We performed two experiments, soliciting respectively 2300, and 900 MTurk 
workers to label images depicting food, kitchen- and dinnerware, and cultural
scenes.  Both experiments had four treatments, and in all treatments of a 
given experiment, the last five ``test'' tasks performed by the workers were 
the same.  Two treatments for each experiment were used to explore inter-task 
effects, by making the initial five tasks for these treatments different.  
For example, in the first 
experiment, the test images depicted clutterd scenes of food and objects,
while ``initial'' images depicted either food only, or non-food objects 
(see Fig. 1).  Although we divide the tasks into initial tasks and test tasks,
no distinction is apparent from the perspective of the worker. Another two 
treatments in both experiments explored framing, 
by presenting the worker with language suggesting a particular research 
intent.  For example, in the first experiment, we stated 
``funded by the laborotary for the visual perception of \{ Objects and Tools 
$\vert$ Food and Ingredients \}''.  For emphasis, the frame text was presented
free of surrounding text, with the variable text in boldface and larger font.

The experiments and treatments are 
summarized in Table 1, which also introduces short names which we will use to 
identify specific experimental treatments and groupings.
An additional difference between \texttt{exp1} and \texttt{exp2} is that,
in both treatments for \texttt{task1}, workers were shown the test tasks in
one of five permutations, which allows us to analyze the dependance of 
inter-task effects on task proximity, independant of task content.
We provide a fuller account of the experimental design, and exhibit all the
task materials used, in the supplementary material.

\setlength{\tabcolsep}{3pt}
\begin{table}[t]
\centering
\begin{tabular}{ c c c c c }
		\hline \noalign{\smallskip}
		\multicolumn{3}{c}{\textbf{Treatment name}} & \parbox[c]{1.6cm}{\centering \textbf{Framing\\ Prime}} & \parbox[c]{1.7cm}{\centering \textbf{Initial\\ image set}}	\\

		\noalign{\smallskip} \hline \noalign{\smallskip}

		\multirow{6}{*}{ \parbox[c]{1cm}{ \phantom{XXX} exp1}} 
			& \multirow{2}{*}{task1} & food & none & food\\
			& & obj & none & objects\\

			\noalign{\smallskip} \cline{2-5} \noalign{\smallskip}
			& \multirow{2}{*}{frame1} & food & food & none\\
			& & obj & objects & none\\

			\noalign{\smallskip} \cline{2-5} \noalign{\smallskip}
			& \multirow{2}{*}{echo} & food & food & none\\
			& & obj & objects & none\\

		\noalign{\smallskip} \hline \noalign{\smallskip}

		\multirow{4}{*}{exp2} 
			& \multirow{2}{*}{task2}  &  food & none & food\\
			& 	&  cult & none & culture\\
			\noalign{\smallskip} \cline{2-5} \noalign{\smallskip}
			& \multirow{2}{*}{frame2} & food & food & food\\
			& 	& cult & culture & food\\

		\noalign{\smallskip} \hline  
	\end{tabular}

	\caption{ \footnotesize{ 
		We use the treatment names shown to refer to specific sections of 
		the data collected.  For example, \texttt{task2:food} refers to the
		treatment in the second experiment which was not exposed to a frame, 
		and whose initial tasks contained images of food (7th row).
	}}
	\label{table:1}
\end{table}

@@

\paragraph{The strength of intertask effects.}

We assessed the strength of intertask and framing effects
by training a multinomial naive Bayes classifier to predict a worker's 
treatment (from two possibilities) based on the labels they provided 
in the test tasks (see Fig.~\ref{fig:theta}).
We found that in \texttt{task1}, whether a worker had labeled food or non-food 
objects during initial tasks, had a strong effect on their subsequent labels, 
biasing them by more than 30\%.  In comparison, the framing effect in \texttt{frame1} did 
not influence workers' labels enough for the classifier predict their treatment
with an accuracy significantly better than chance.  The results for 
\texttt{exp2} show a similar, but even more pronounced trend, with the bias
induced by intertask effects in \texttt{task2} probably exceeding 40\%.

It is worth noting that \texttt{task1} and \texttt{task2} differ in the kind
of comparison that they make.  In \texttt{task1},
the prime concepts (``food'' and ``objects'') are, in certain ways, very 
similar: both are relatively concrete, well-defined concepts, corresponding
to bounded physical entities that can be directly photographed.
But in \texttt{task2}, the prime concepts (``food'' and ``culture''), 
are very different:  culture is far more abstract and difficult to define
than food, and is not a bounded physical entity.  We will come back to this 
point, but note 
here that these pairs of prime concepts both lead to strong intertask effects
despite exploring very different conceptual dimensions.

As another point of comparison, we included additional framing-like 
treatments in \texttt{exp1} in which we attempted to increase the effect.
In the \texttt{echo} treatment pair, we used a more strongly worded frame:
``The purpose of this study is to understand the visual perception of
\{ Objects and Tools $\vert$ Food and Ingredients \}'', and then worker was
asked to reiterate back the purpose of the study, using a combo-box input.
We expected this to induce a strong effect because it signals that this detail
is of particular concern to the task requester, and more important than any of 
the instructions given beforehand.  
The overall effect of these primes was on par with the intertask 
effects in \texttt{task1} (see Fig.~\ref{fig:theta}), which shows just how
powerful intertask effects are.

\begin{figure}
	\centering
	\includegraphics[scale=1]{figs/theta.pdf}
	\caption{
		Difference in priming observed for various pairs of worker populations.
		The pairs of populations differed in terms of the initial tasks they
		performed (bars labeled by \textit{inter-t.}) or in terms of the framing
		to which they were exposed (bars labeled by \textit{frame}; see table 1 for
		details about the treatments).  The difference in priming was 
		measured based on the accuracy of a naive Bayes classifier in 
		resolving which treatment workers belonged to, as described in the 
		text.  
	}
	\label{fig:theta}
\end{figure}

\paragraph{Dynamics of intertask effects.} Following the observation that
intertask effects are strong, it is natural to ask how the effect of an early
task wanes as the worker proceeds through subsequent ones.  To investigate
this, in \texttt{task1} workers performed the test tasks in one of five 
permutations.  This allows us to average the intertask effect accross different
test tasks when they occupy the same position.  We found that the effect was
strongest for the first test image, and that a significant effect persisted
through the five test tasks (see Fig.~\ref{fig:theta}B).  This shows the 
extent to which the first 
test task re-primes workers, washing out the effect of initial tasks, but that
the effect is quite persistent.  In future work, it would be interesting to
investigate whether the first few tasks have a particularly strong priming
effect because they ``set the tone'' for the batch.

This set-up also allows us to see how much the each test tasks differ in their 
susceptibility to intertask effects, independant of their position among the
others.  There does appear to be substantial differences, but relative
position of the test task is more important (Fig.~\ref{fig:theta}C).

\paragraph{Effects on workers' direction of focus} 
\begin{figure}
	\centering
	\includegraphics[scale=1]{figs/delta_food.pdf}
	\caption{
		Difference in priming observed for various pairs of worker populations.
		The pairs of populations differed in terms of the initial tasks they
		performed (bars labeled by \textit{inter-t.}) or in terms of the framing
		to which they were exposed (bars labeled by \textit{frame}; see table 1 for
		details about the treatments).  The difference in priming was 
		measured based on the accuracy of a naive Bayes classifier in 
		resolving which treatment workers belonged to, as described in the 
		text.  
	}
	\label{fig:food}
\end{figure}

It is reasonable to expect that exposing a worker to tasks containing images
of food would increase her propensity to provide food-related labels in 
subsequent tasks.  To test whether this straightforward elicitation occured,
we used the wordnet corpus to operationalize the notion of food-related 
labels.  While the coverage of most western food-related words is quite good
we found that the names of many ethnic foods are missing, so we augmented
the corpus by crawling the world food section of the recipes website 
allrecipes.com.  We discuss the validation of our operationalization of 
food-words in detail in the supplementary material.

For \texttt{task1} and \texttt{frame1} we did find that exposing workers to 
food-oriented primes led them to increase their provision of food-related 
labels by about 2-3\% (see Fig.~\ref{fig:food}).  
In \texttt{echo} this effect was stronger
\texttt{echo:food} provided about 8\% more food-related words than 
\texttt{echo:obj}.  Indeed, the \%-increase in food labels observed for
\texttt{echo} was significantly larger than that for \texttt{taks1}, which is 
interesting
given that the overall priming effect observed for \texttt{task1} and 
\texttt{echo} were on par (c.f. Fig.~\ref{fig:theta}).
This makes sense, given that \texttt{echo:food} explicitly calls for a focus 
on food, whereas \texttt{task1:food} probably works by activating certain 
memories and concepts, with a more complex impact on labels.
